CODECHECK: An open-science initiative for the independent execution of computations underlying research articles

### 2021-01-14 codecheck-paper.tex
trust vs. transparency

Note that the task of the CODECHECKER is to check the ``mechanics'' of the
workflow, not to evaluate its correctness. 
Stodden et~al. \cite{stodden_setting_2013} distinguish between \emph{verification}
and \emph{validation}
in the context of Mathematics. Following their definition, a CODECHECK
ensures a verification\footnote{Note this is a different use as the
\emph{Verification Reports} 
article type in the journal \emph{Cortex} \cite{chambers_verification_2020}},
of computational results, i.e.,~checking that the code correctly produces the 
output it claims to create, but not a validation, i.e.,~does the code implement
the right algorithm to solve the problem that is researched.
Nevertheless, the simple attempt to reproduce may highlight shortcomings in
a submission with respect to transparency as it may be required by a journal's
policy and help to resolve misunderstandings and discrepancies
(cf.~\cite{christian_journal_2020}).

@@

\textbf{5. Open by default and transitional by disposition.} % or "design"?
By default, unless there are strong
reasons to the contrary (e.g.,~clinically-sensitive data), all code and
data, both from author and codechecker, will be made freely available, 
i.e.,~under open licenses, before or when the certificate is published.
Note that openness is not required for the paper itself as to accommodate 
journals to transition to sustainable Open Access models.
The code and data publication should follow community good practices.
By disposition, a CODECHECK workflow should be designed (a) as a recognition of
shortcomings in asserting reproducibility of submissions and (b)
with an understanding of the
complex situation that lead to this. However, establishing a CODECHECK workflow
should go hand in hand with initiatives and a perspective to improve education
around reproducibility of methods and communication of comprehensive 
computational workflows, with the long term goal for a specific community or
journal's target audience to reach a state where the level of verification 
and trust provided by codecheckers becomes part of peer review and regular
merit of all accepted papers.

@@

CODECHECK works!
We have created a considerable number of certificates to demonstrate it.
The creation of the certificates and the interactions with authors and
editors not only improved the principles and the workflow, but also
confirmed the approach taken. This result corroborates findings from
similar evaluations of reproducible computational research in journals and
at conferences.
CODECHECKs can increase trust in scientific results.
The set of 
shared principles and common name, through recognition value,
allow researchers
across publication venues to judge the level of scrutiny that results have
faced. CODECHECK requires direct acknowledgement of the contribution to the
scientific community in the role of codechecker, not indirectly via citations 
of reproductions or informal crediation and closed inner circles.
Codechecking can be a worthwhile educational activity, especially for
ECRs.

>>>>

Note that the task of the CODECHECKER is to check the ``mechanics'' of the
workflow, not to evaluate its correctness. 
Stodden et~al. \cite{stodden_setting_2013} distinguish between \emph{verification}
and \emph{validation}
in the context of Mathematics. Following their definition, a CODECHECK
ensures a verification\footnote{Note this is a different use as the
\emph{Verification Reports} 
article type in the journal \emph{Cortex} \cite{chambers_verification_2020}},
of computational results, i.e.,~checking that the code correctly produces the 
output it claims to create, but not a validation, i.e.,~does the code implement
the right algorithm to solve the problem that is researched.
Nevertheless, the simple attempt to reproduce may highlight shortcomings in
a submission to meet journal policy requirements
(cf.~\cite{christian_journal_2020})
and can effectively increase transparency and thereby improve practices
(cf.~\cite{nosek_scientific_2012})
even if the check goes not into all details.

@@

\textbf{5. Open by default and transitional by disposition.} % or "design"?
By default, unless there are strong
reasons to the contrary (e.g.,~clinically-sensitive data), all code and
data, both from author and codechecker, will be made freely available, 
i.e.,~under open licenses, before or when the certificate is published.
Note that openness is not required for the paper itself as to accommodate 
journals to transition to sustainable Open Access models.
The code and data publication should follow community good practices.
By disposition, a CODECHECK workflow should be designed (a) as a recognition of
shortcomings in asserting reproducibility of submissions and (b)
with an understanding of the
complex situation that lead to this. However, establishing a CODECHECK workflow
should go hand in hand with initiatives and a perspective to improve education
around reproducibility of methods and communication of comprehensive 
computational workflows, with the long term goal for a specific community or
journal's target audience to reach a state where the level of verification 
ensured by codecheckers becomes part of peer review and regular
merit of all accepted papers.

@@

CODECHECK works!
We have created a considerable number of certificates to demonstrate it.
The creation of the certificates and the interactions with authors and
editors not only improved the principles and the workflow, but also
confirmed the approach taken. This result corroborates findings from
similar evaluations of reproducible computational research in journals and
at conferences.
CODECHECKs increase transparency of the checked papers and can contribute
to building trust in scientific results.
% "science earns trust through transparency" > https://twitter.com/briannosek/status/1290240244101521408
The set of 
shared principles and common name, through recognition value,
allow researchers
across publication venues to judge the level of scrutiny that results have
faced. CODECHECK requires direct acknowledgement of the contribution to the
scientific community in the role of codechecker, not indirectly via citations 
of reproductions or informal crediation and closed inner circles.
Codechecking can be a worthwhile educational activity, especially for
ECRs.
